\textbf{SPECFEM3D}\\ \textbf{Cartesian}\\ \textbf{User Manual}

### 2023-11-15 doc/USER_MANUAL/03_mesh_generation.tex

For an elastic material:
\begin{description}
\item [{material\_id}] An integer value which is unique for this material.
\item [{Vp}] P-wave speed of the material (given in m/s).
\item [{Vs}] S-wave speed of the material (given in m/s).
\item [{rho}] density of the material (given in kg/m$^{3}$).
\item [{Q}] quality factor to use in case of a simulation with attenuation
turned on. It should be between 1 and 9000. In case no attenuation
information is available, it can be set to zero. You can either specify a single Q value,
in which case it will be assumed to be pure shear attenuation $Q_{\mu}$, or two
separate values for bulk and shear attenuation, $Q_{\kappa}$ and $Q_{\mu}$ respectively.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other
see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and\newline
\texttt{utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}.\newline

@@

To convert one to the other see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and the helper code\newline
\texttt{conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90} in folder
\texttt{utils/attenuation/}.

@@

Input for the mesh generation program is provided through the parameter
file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
(To see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/meshfem3D\_examples/}.)
Before running the mesher, a number of parameters need to be set in
the \texttt{Mesh\_Par\_file}. This requires a basic understanding
of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
and \citet{KoLiTrSuStSh04}.\newline

@@

Depending on your system, please check how to run executables with MPI.
To run the mesher on your cluster, we provide some example scripts for you to modify accordingly in \texttt{utils/Cluster/}.
For some installations, you might need to provide a file
that tells MPI what compute nodes to use for the simulations. In this case, the
file must have a number of entries (one entry per line) at least equal
to the number of processors needed for the run.
A sample file name is \texttt{mymachines}. This file is not used by the mesher
or solver, but is required by the \texttt{go\_mesher***} and \texttt{go\_solver**}
default job submission scripts provided in directory \texttt{utils/Cluster/}.
To run the mesher, it is most easily accomplished based upon the \texttt{go\_mesher***} script.
When you run on a PC cluster, the script assumes that the nodes are
named \texttt{n001}, \texttt{n002}, etc.
If this is not the case, change the \texttt{tr -d `n'} line in the script.
You may also need to edit the last command
at the end of the script that invokes the \texttt{mpirun} command.
See Chapter~\ref{cha:Scheduler}
for information about running the code on a system with a scheduler,
e.g., LSF.\newline

@@

The way we prescribe material property for porous material in SPECFEM3D depends on wheter the internal mesher \texttt{xmeshfem3D}
or an external mesher like CUBIT is taken.
For external meshers which provide the resulting meshes in a folder like \texttt{MESH/},
we use a file \texttt{nummaterial\_poroelastic\_file}, which needs to be added to this directory \texttt{MESH/}.
The line format for poroelastic materials is as follow:
{\small \begin{verbatim}
  rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr
\end{verbatim}}
\begin{description}
\item where\newline
\texttt{rho\_s} = solid density,\newline
\texttt{rho\_f} = fluid density,\newline
\texttt{phi} = porosity,\newline
\texttt{tort} = tortuosity,\newline
\texttt{kxx} = xx component of permeability tensor,\newline
\texttt{kxy} = xy,yx components of permeability tensor,\newline
\texttt{kyy} = yy component of permeability tensor,\newline
\texttt{kxz} = xz,zx components of permeability tensor,\newline
\texttt{kzz} = zz component of permeability tensor,\newline
\texttt{kappa\_s} = solid bulk modulus,\newline
\texttt{kappa\_f} = fluid bulk modulus,\newline
\texttt{kappa\_fr} = frame bulk modulus,\newline
\texttt{eta\_f} = fluid viscosity,\newline
\texttt{mu\_fr} = frame shear modulus.\newline
\end{description}
Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: EXAMPLES/homogeneous\_poroelastic).\newline

>>>>

For an elastic material:
\begin{description}
\item [{material\_id}] An integer value which is unique for this material.
\item [{Vp}] P-wave speed of the material (given in m/s).
\item [{Vs}] S-wave speed of the material (given in m/s).
\item [{rho}] density of the material (given in kg/m$^{3}$).
\item [{Q}] quality factor to use in case of a simulation with attenuation
turned on. It should be between 1 and 9000. In case no attenuation
information is available, it can be set to zero. You can either specify a single Q value,
in which case it will be assumed to be pure shear attenuation $Q_{\mu}$, or two
separate values for bulk and shear attenuation, $Q_{\kappa}$ and $Q_{\mu}$ respectively.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other
see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and\newline
\texttt{utils/small\_utilities/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}.\newline

@@

To convert one to the other see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and the helper code\newline
\texttt{conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90} in folder
\texttt{utils/small\_utilities/attenuation/}.

@@

Input for the mesh generation program is provided through the parameter
file \texttt{Mesh\_Par\_file}, which resides in the subdirectory \texttt{DATA/meshfem3D\_files/}.
(To see how to use it, see the EXAMPLES specific to the internal mesher in directory \texttt{EXAMPLES/applications/meshfem3D\_examples/}.)
Before running the mesher, a number of parameters need to be set in
the \texttt{Mesh\_Par\_file}. This requires a basic understanding
of how the SEM is implemented, and we encourage you to read \citet{KoVi98,KoTr99}
and \citet{KoLiTrSuStSh04}.\newline

@@

Depending on your system, please check how to run executables with MPI.
To run the mesher on your cluster, we provide some example scripts for you to modify accordingly in \texttt{utils/scripts/Cluster/}.
For some installations, you might need to provide a file
that tells MPI what compute nodes to use for the simulations. In this case, the
file must have a number of entries (one entry per line) at least equal
to the number of processors needed for the run.
A sample file name is \texttt{mymachines}. This file is not used by the mesher
or solver, but is required by the \texttt{go\_mesher***} and \texttt{go\_solver**}
default job submission scripts provided in directory \texttt{utils/scripts/Cluster/}.
To run the mesher, it is most easily accomplished based upon the \texttt{go\_mesher***} script.
When you run on a PC cluster, the script assumes that the nodes are
named \texttt{n001}, \texttt{n002}, etc.
If this is not the case, change the \texttt{tr -d `n'} line in the script.
You may also need to edit the last command
at the end of the script that invokes the \texttt{mpirun} command.
See Chapter~\ref{cha:Scheduler}
for information about running the code on a system with a scheduler,
e.g., LSF.\newline

@@

The way we prescribe material property for porous material in SPECFEM3D depends on wheter the internal mesher \texttt{xmeshfem3D}
or an external mesher like CUBIT is taken.
For external meshers which provide the resulting meshes in a folder like \texttt{MESH/},
we use a file \texttt{nummaterial\_poroelastic\_file}, which needs to be added to this directory \texttt{MESH/}.
The line format for poroelastic materials is as follow:
{\small \begin{verbatim}
  rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr
\end{verbatim}}
\begin{description}
\item where\newline
\texttt{rho\_s} = solid density,\newline
\texttt{rho\_f} = fluid density,\newline
\texttt{phi} = porosity,\newline
\texttt{tort} = tortuosity,\newline
\texttt{kxx} = xx component of permeability tensor,\newline
\texttt{kxy} = xy,yx components of permeability tensor,\newline
\texttt{kyy} = yy component of permeability tensor,\newline
\texttt{kxz} = xz,zx components of permeability tensor,\newline
\texttt{kzz} = zz component of permeability tensor,\newline
\texttt{kappa\_s} = solid bulk modulus,\newline
\texttt{kappa\_f} = fluid bulk modulus,\newline
\texttt{kappa\_fr} = frame bulk modulus,\newline
\texttt{eta\_f} = fluid viscosity,\newline
\texttt{mu\_fr} = frame shear modulus.\newline
\end{description}
Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: \texttt{EXAMPLES/applications/homogeneous\_poroelastic}).\newline

### 2023-11-15 doc/USER_MANUAL/07_adjoint_simulations.tex

\begin{enumerate}
\item First, run a regular forward simulation (\texttt{SIMULATION\_TYPE =
1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
script:}{\small \par}
\begin{verbatim}
utils/change_simulation_type.pl -f
\end{verbatim}

@@

\begin{itemize}
\item Depending on what type of misfit function is used for the source inversion,
adjoint sources need to be computed from the original recorded seismograms
for the selected stations and saved in a sub-directory called \texttt{SEM/}
in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
file, and \texttt{BX?} represents the component name of a particular
adjoint seismogram. Please note that the band code can change depending
on your sampling rate (see Appendix~\ref{cha:channel-codes} for
further details).
\item The adjoint seismograms are in the same format as the original seismogram
(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
and record length.
\end{itemize}
\item Notice that even if you choose to time reverse only one component
from one specific station, you still need to supply all three components
because the code is expecting them (you can set the other two components
to be zero).
\item Also note that since time-reversal is done in the code itself, no
explicit time-reversing is needed for the preparation of the adjoint
sources, i.e., the adjoint sources are in the same forward time sense
as the original recorded seismograms.
\end{enumerate}
\item \textbf{Set the related parameters and run the adjoint simulation}\newline
In the \texttt{DATA/Par\_file}, set the two related parameters to
be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
to modify the \texttt{DATA/Par\_file} automatically (\texttt{change\_simulation\_type.pl
-a}). Then run the solver to launch the adjoint simulation.
\item \textbf{Collect the seismograms at the original source location}

>>>>

\begin{enumerate}
\item First, run a regular forward simulation (\texttt{SIMULATION\_TYPE =
1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
these two variables using the \texttt{\small utils/scripts/change\_simulation\_type.pl}{\small{}
script:}{\small \par}
\begin{verbatim}
utils/scripts/change_simulation_type.pl -f
\end{verbatim}

@@

\begin{itemize}
\item Depending on what type of misfit function is used for the source inversion,
adjoint sources need to be computed from the original recorded seismograms
for the selected stations and saved in a sub-directory called \texttt{SEM/}
in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
file, and \texttt{BX?} represents the component name of a particular
adjoint seismogram. Please note that the band code can change depending
on your sampling rate (see Appendix~\ref{cha:channel-codes} for
further details).
\item The adjoint seismograms are in the same format as the original seismogram
(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
and record length.
\end{itemize}
\item Notice that even if you choose to time reverse only one component
from one specific station, you still need to supply all three components
because the code is expecting them (you can set the other two components
to be zero).
\item Also note that since time-reversal is done in the code itself, no
explicit time-reversing is needed for the preparation of the adjoint
sources, i.e., the adjoint sources are in the same forward time sense
as the original recorded seismograms.
\end{enumerate}
\item \textbf{Set the related parameters and run the adjoint simulation}\newline
In the \texttt{DATA/Par\_file}, set the two related parameters to
be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
More conveniently, use the scripts \texttt{utils/scripts/change\_simulation\_type.pl}
to modify the \texttt{DATA/Par\_file} automatically (\texttt{change\_simulation\_type.pl
-a}). Then run the solver to launch the adjoint simulation.
\item \textbf{Collect the seismograms at the original source location}

### 2023-11-15 doc/USER_MANUAL/11_graphics.tex
- added a new parameter "MOVIE_VOLUME_STRESS" in the Par_file to save the stress tensor components in the entire volume for the viscoelastic simulatons.
adds support for MOVIE_VOLUME_STRESS flag and output

To make a surface or volume movie of the simulation, set parameters
\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{DATA/Par\_file}. Turning
on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
large output files. \texttt{MOVIE\_VOLUME} files are saved in the
\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
set, otherwise the velocity field is saved. The look of a movie is
determined by the half-duration of the source. The half-duration should
be large enough so that the movie does not contain frequencies that
are not resolved by the mesh, i.e., it should not contain numerical
noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
of each source in the \texttt{CMTSOLUTION} file is replaced by
\begin{quote}
\[
\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
\]
\end{quote}

@@

You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located. As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:

{\footnotesize
\begin{verbatim}
./bin/xcombine_vol_data 0 3 velocity_Z_it000400 ./OUTPUT_FILES/DATABASES_MPI ./OUTPUT_FILES 0
\end{verbatim}
}
Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).\newline

@@

\noindent
For this task, you can use and modify the
script \texttt{mesh2vtu.pl} located in directory \texttt{\small utils/Visualization/Paraview}, for example:

{\footnotesize
\begin{verbatim}
mesh2vtu.pl -i velocity_Z_it000400.mesh -o velocity_Z_it000400.vtu
\end{verbatim}
}

@@

A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
can help to figure out the slice numbers that lie along the great
circle path. It applies to meshes created with the internal mesher
\texttt{xmeshfem3D}.
\begin{enumerate}
\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
and \texttt{output\_solver} files, and run:

{\small
\begin{verbatim}
slice_number.pl Mesh_Par_file output_solver.txt slice_file
\end{verbatim}
}
which will generate a \texttt{slices\_file}.

@@

After obtaining the slice files, you can collect the corresponding
kernel files from the given slices.
\begin{enumerate}
\item You can use or modify the script \texttt{utils/copy\_basin\_database.pl}
to accomplish this:

{\small
\begin{verbatim}
utils/copy_database.pl slice_file lsf_machine_file filename [jobid]
\end{verbatim}
}
where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
generated by the LSF scheduler, }\texttt{\small filename}{\small{}
is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
where all the kernel files are stored.}{\small \par}

@@

\begin{enumerate}
\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
grid file) format which can be viewed in ParaView. For this task,
you can use and modify the script \texttt{mesh2vtu.pl} located in
directory \texttt{utils/Visualization/Paraview/}, for example:

{\footnotesize
\begin{verbatim}
mesh2vtu.pl -i file.mesh -o file.vtu
\end{verbatim}
}

>>>>

To make a surface or volume movie of the simulation, set parameters
\texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME}, \texttt{MOVIE\_TYPE},
and \texttt{NTSTEP\_BETWEEN\_FRAMES} in the \texttt{DATA/Par\_file}. Turning
on the movie flags, in particular \texttt{MOVIE\_VOLUME}, produces
large output files. \texttt{MOVIE\_VOLUME} files are saved in the
\texttt{LOCAL\_PATH} directory, whereas \texttt{MOVIE\_SURFACE} output
files are saved in the \texttt{OUTPUT\_FILES} directory. We save the
displacement field if the parameter \texttt{SAVE\_DISPLACEMENT} is
set, otherwise the velocity field is saved.
Additionally, if the \texttt{MOVIE\_VOLUME\_STRESS} is set to \texttt{.true.},
stress tensor components are saved in the entire volume.\newline

The look of a movie is determined by the half-duration of the source.
The half-duration should be large enough so that the movie does not contain frequencies that
are not resolved by the mesh, i.e., it should not contain numerical
noise. This can be accomplished by selecting a CMT \texttt{HALF\_DURATION}
> 1.1 $\times$ smallest period (see figure \ref{fig:CMTSOLUTION-file}).
When \texttt{MOVIE\_SURFACE} = .\texttt{true.}, the half duration
of each source in the \texttt{CMTSOLUTION} file is replaced by
\begin{quote}
\[
\sqrt{(}\mathrm{\mathtt{HALF\_DURATIO}\mathtt{N}^{2}}+\mathrm{\mathtt{HDUR\_MOVI}\mathtt{E}^{2}})
\]
\end{quote}

@@

You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located. As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:
{\footnotesize
\begin{verbatim}
./bin/xcombine_vol_data 0 3 velocity_Z_it000400 ./OUTPUT_FILES/DATABASES_MPI ./OUTPUT_FILES 0
\end{verbatim}
}
Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).\newline

@@

\noindent
For this task, you can use and modify the
script \texttt{mesh2vtu.pl} located in directory \texttt{\small utils/Visualization/Paraview}, for example:
{\footnotesize
\begin{verbatim}
mesh2vtu.pl -i velocity_Z_it000400.mesh -o velocity_Z_it000400.vtu
\end{verbatim}
}

@@

A Perl script \texttt{slice\_number.pl} located in directory \texttt{utils/Visualization/Paraview/}
can help to figure out the slice numbers that lie along the great
circle path. It applies to meshes created with the internal mesher
\texttt{xmeshfem3D}.
\begin{enumerate}
\item On machines where you have access to the script, copy the \texttt{Mesh\_Par\_file},
and \texttt{output\_solver} files, and run:
{\small
\begin{verbatim}
slice_number.pl Mesh_Par_file output_solver.txt slice_file
\end{verbatim}
}
which will generate a \texttt{slices\_file}.

@@

After obtaining the slice files, you can collect the corresponding
kernel files from the given slices.
\begin{enumerate}
\item You can use or modify the script \texttt{utils/scripts/Cluster/copy\_basin\_database.pl}
to accomplish this:
{\small
\begin{verbatim}
utils/scripts/Cluster/copy_basin_database.pl slice_file lsf_machine_file filename [jobid]
\end{verbatim}
}
where \texttt{\small lsf\_machine\_file}{\small{} is the machine file
generated by the LSF scheduler, }\texttt{\small filename}{\small{}
is the kernel name (e.g., }\texttt{\small rho\_kernel}{\small , }\texttt{\small alpha\_kernel}{\small{}
and }\texttt{\small beta\_kernel}{\small ), and the optional }\texttt{\small jobid}{\small{}
is the name of the subdirectory under }\texttt{\small LOCAL\_PATH}{\small{}
where all the kernel files are stored.}{\small \par}

@@

\begin{enumerate}
\item We next convert the \texttt{.mesh} file into the VTU (Unstructured
grid file) format which can be viewed in ParaView. For this task,
you can use and modify the script \texttt{mesh2vtu.pl} located in
directory \texttt{utils/Visualization/Paraview/}, for example:
{\footnotesize
\begin{verbatim}
mesh2vtu.pl -i file.mesh -o file.vtu
\end{verbatim}
}

### 2023-11-15 doc/USER_MANUAL/12_running_scheduler.tex

The code is usually run on large parallel machines, often PC clusters,
most of which use schedulers, i.e., queuing or batch management systems
to manage the running of jobs from a large number of users. The following
considerations need to be taken into account when running on a system
that uses a scheduler:
\begin{itemize}
\item The processors/nodes to be used for each run are assigned dynamically
by the scheduler, based on availability. Therefore, in order for the
\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables
(or between successive runs of the solver) to have access to the same
database files (if they are stored on hard drives local to the nodes
on which the code is run), they must be launched in sequence as a
single job.
\item On some systems, the nodes to which running jobs are assigned are
not configured for compilation. It may therefore be necessary to pre-compile
both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D}
executables.
\item One feature of schedulers/queuing systems is that they allow submission
of multiple jobs in a ``launch and forget'' mode. In order to take
advantage of this property, care needs to be taken that output and
intermediate files from separate jobs do not overwrite each other,
or otherwise interfere with other running jobs.
\end{itemize}
Examples of job scripts can be found in the \texttt{\small utils/Cluster/}{\small{}
directory and can straightforwardly be modified and adapted to meet
more specific running needs.}{\small \par}

>>>>

The code is usually run on large parallel machines, often PC clusters,
most of which use schedulers, i.e., queuing or batch management systems
to manage the running of jobs from a large number of users. The following
considerations need to be taken into account when running on a system
that uses a scheduler:
\begin{itemize}
\item The processors/nodes to be used for each run are assigned dynamically
by the scheduler, based on availability. Therefore, in order for the
\texttt{xgenerate\_databases} and the \texttt{xspecfem3D} executables
(or between successive runs of the solver) to have access to the same
database files (if they are stored on hard drives local to the nodes
on which the code is run), they must be launched in sequence as a
single job.
\item On some systems, the nodes to which running jobs are assigned are
not configured for compilation. It may therefore be necessary to pre-compile
both the \texttt{xgenerate\_databases} and the \texttt{xspecfem3D}
executables.
\item One feature of schedulers/queuing systems is that they allow submission
of multiple jobs in a ``launch and forget'' mode. In order to take
advantage of this property, care needs to be taken that output and
intermediate files from separate jobs do not overwrite each other,
or otherwise interfere with other running jobs.
\end{itemize}
Examples of job scripts can be found in the \texttt{\small utils/scripts/Cluster/}{\small{}
directory and can straightforwardly be modified and adapted to meet
more specific running needs.}{\small \par}

### 2023-11-15 doc/USER_MANUAL/13_changing_the_model.tex

\noindent
Each data record line provides the velocity model values in a format like:
{\small
\begin{verbatim}
# data record format: purely isotropic model
x y z   vp    vs    rho
..
\end{verbatim}
}
\noindent
or
{\small
\begin{verbatim}
# data record format: anelastic isotropic model
x y z   vp    vs    rho Qp  Qs
..
\end{verbatim}
}
\noindent
where  \texttt{x}, \texttt{y}, \texttt{z} are the grid point position, \texttt{vp}, \texttt{vs} and \texttt{rho} the P- and S-wave speeds and density, and \texttt{Qp} and \texttt{Qs} the quality factors for P- and S-wave speeds. The quality factors are optional and can be omitted for purely elastic models (the tomography routine will recognize both formats). Internally, the quality factors will be converted to bulk and shear attenuation values,
$Q_{\kappa}$ and $Q_{\mu}$ respectively \citep{AnHa78}.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
For simulations with attenuation, please note that the Vp- and Vs-velocities of your model are given for a reference
frequency. To change this reference frequency, you change the value
of \texttt{ATTENUATION\_f0\_REFERENCE} defined (in Hz) in input file \texttt{DATA/Par\_file}.\newline

@@

To use your own external model, you can set in the \texttt{DATA/Par\_file}
the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
and/or (an)elastic (attenuation) models may then be superimposed onto
the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
located in directory \texttt{src/generate\_databases}. The call to
this routine would be as follows
\begin{verbatim}
call model_external_values(xmesh, ymesh, zmesh, &
            rho, vp, vs, qkappa_atten, qmu_atten, iflag_aniso, idomain_id)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
\end{description}
Output to this routine consists of:
\begin{description}
\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
\item [{\texttt{qkappa\_atten,qmu\_atten}}] Bulk and shear wave quality factor: $0<Q_{\kappa,\mu}<9000$.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.
\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
no anisotropy or $1$ using anisotropic model parameters as defined
in routine file \texttt{model\_aniso.f90}
\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
for elastic, $3$ for poroelastic materials.
\end{description}
Note that the resolution and maximum value of anelastic models are
truncated. This speeds the construction of the standard linear solids
during the meshing stage. To change the resolution, currently at one
significant figure following the decimal, or the maximum value (9000),
consult \texttt{setup/constants.h}.

@@

An example demonstrating the use SEP models is given in:
\texttt{EXAMPLES/meshfem3D\_examples/sep\_bathymetry}.

>>>>

\noindent
Each data record line provides the velocity model values in a format like:
{\small
\begin{verbatim}
# data record format: purely isotropic model
x y z   vp    vs    rho
..
\end{verbatim}
}
\noindent
or
{\small
\begin{verbatim}
# data record format: anelastic isotropic model
x y z   vp    vs    rho Qp  Qs
..
\end{verbatim}
}
\noindent
where  \texttt{x}, \texttt{y}, \texttt{z} are the grid point position, \texttt{vp}, \texttt{vs} and \texttt{rho} the P- and S-wave speeds and density, and \texttt{Qp} and \texttt{Qs} the quality factors for P- and S-wave speeds. The quality factors are optional and can be omitted for purely elastic models (the tomography routine will recognize both formats). Internally, the quality factors will be converted to bulk and shear attenuation values,
$Q_{\kappa}$ and $Q_{\mu}$ respectively \citep{AnHa78}.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and \texttt{utils/small\_utilities/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}.
For simulations with attenuation, please note that the Vp- and Vs-velocities of your model are given for a reference
frequency. To change this reference frequency, you change the value
of \texttt{ATTENUATION\_f0\_REFERENCE} defined (in Hz) in input file \texttt{DATA/Par\_file}.\newline

@@

To use your own external model, you can set in the \texttt{DATA/Par\_file}
the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
and/or (an)elastic (attenuation) models may then be superimposed onto
the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
located in directory \texttt{src/generate\_databases}. The call to
this routine would be as follows
\begin{verbatim}
call model_external_values(xmesh, ymesh, zmesh, &
            rho, vp, vs, qkappa_atten, qmu_atten, iflag_aniso, idomain_id)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
\end{description}
Output to this routine consists of:
\begin{description}
\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
\item [{\texttt{qkappa\_atten,qmu\_atten}}] Bulk and shear wave quality factor: $0<Q_{\kappa,\mu}<9000$.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and \texttt{utils/small\_utilities/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}.
\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
no anisotropy or $1$ using anisotropic model parameters as defined
in routine file \texttt{model\_aniso.f90}
\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
for elastic, $3$ for poroelastic materials.
\end{description}
Note that the resolution and maximum value of anelastic models are
truncated. This speeds the construction of the standard linear solids
during the meshing stage. To change the resolution, currently at one
significant figure following the decimal, or the maximum value (9000),
consult \texttt{setup/constants.h}.

@@

An example demonstrating the use SEP models is given in:
\texttt{EXAMPLES/applications/meshfem3D\_examples/sep\_bathymetry}.

### 2023-11-15 doc/USER_MANUAL/C_troubleshooting.tex

\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
at the skewness table printed in the \texttt{output\_mesher.txt} file
after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:

@@

To give you an idea why some of the elements are distorted, see the
following figure \ref{fig:mesh.vp} of the mesh you obtain in the
example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.9\textwidth]{figures/mesh_vp.jpg}
\par\end{centering}

\caption{Paraview visualization using the mesh vtk-files for the example given
in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}

>>>>

\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
at the skewness table printed in the \texttt{output\_mesher.txt} file
after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/applications/meshfem3D\_examples/simple\_model/}:

@@

To give you an idea why some of the elements are distorted, see the
following figure \ref{fig:mesh.vp} of the mesh you obtain in the
example \texttt{EXAMPLES/applications/meshfem3D\_examples/simple\_model/}.
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.9\textwidth]{figures/mesh_vp.jpg}
\par\end{centering}

\caption{Paraview visualization using the mesh vtk-files for the example given
in \texttt{EXAMPLES/applications/meshfem3D\_examples/simple\_model/}.}
